Inference of proto-neutron star properties from gravitational-wave data in core-collapse supernovae

### 2020-11-10 conclusion.tex
prep version to be circulated in LVK

We have investigated the performance of the algorithm in the case of an optimally oriented source detected by a singe GW detector. Our numerical signals have been injected into 100 Gaussian noise realisations whose PSD follow the spectra of the different GW detectors analyzed. We have found that for Advanced LIGO and Advanced Virgo, the ratio $r$ can be reconstructed with a good accuracy in the case of a galactic CCSN (i.e.~for distances of ${\cal O}(10\, {\rm kpc})$). This holds for a wide range of progenitor masses, the quality of the inference mainly depending on the signal-to-noise ratio of the event. For third-generation GW detectors such as the Einstein Telescope and the Cosmic Explorer, however, we obtain an order of magnitude improvement, as the $\mbox{}^2g_2$ ratio can be reconstructed for sources at distances of ${\cal O}(100\, {\rm kpc})$. In particular, Cosmic Explorer in its stage 2 configuration yields the best performance for all waveforms we have considered thanks to its excellent sensitivity in the \unit[100-1000]{Hz} range. Among the three configurations of the Einstein Telescope, ET-D provides the best performance, especially for our set waveforms with the highest progenitor masses (25 $\Msol$ and 40 $\Msol$). Comparing the estimated distances for ET-B and the other third-generations detectors, having a good sensitivity below \unit[200]{Hz} seems the most important factor to detect high mass progenitor signals.

>>>>

We have investigated the performance of the algorithm in the case of an optimally oriented source detected by a singe GW detector. Our numerical signals have been injected into 100 Gaussian noise realisations whose PSD follow the spectra of the different GW detectors analyzed. We have found that for Advanced LIGO and Advanced Virgo, the ratio $r$ can be reconstructed with a good accuracy in the case of a galactic CCSN (i.e.~for distances of ${\cal O}(10\, {\rm kpc})$). This holds for a wide range of progenitor masses, the quality of the inference mainly depending on the signal-to-noise ratio of the event. For third-generation GW detectors such as the Einstein Telescope and the Cosmic Explorer, however, we obtain an order of magnitude improvement, as the $\mbox{}^2g_2$ ratio can be reconstructed for sources at distances of ${\cal O}(\unit[100]{kpc})$. In particular, Cosmic Explorer in its stage 2 configuration yields the best performance for all waveforms we have considered thanks to its excellent sensitivity in the \unit[100-1000]{Hz} range. Among the three configurations of the Einstein Telescope, ET-D provides the best performance, especially for our set waveforms with the highest progenitor masses (25 $\Msol$ and 40 $\Msol$). Comparing the estimated distances for ET-B and the other third-generations detectors, having a good sensitivity below \unit[200]{Hz} seems the most important factor to detect high mass progenitor signals.

### 2020-11-10 introduction.tex
prep version to be circulated in LVK

The life of sufficiently massive stars, i.e.~those born with masses between $\sim 8$~M$_\odot$ and $\sim 120$~M$_\odot$, ends with the collapse of {the} iron core under {its} own gravity, leading {to} the formation of a neutron star {(NS)} or a black hole (BH), {and} followed (typically but not necessarily in the BH case) by {a supernova} explosion. Nearby core-collapse supernova (CCSN) explosions are expected to be sources of gravitational waves (GWs) and they could be 
%are one of the main candidates for 
the next great discovery of current ground-based observatories. However, these are relative rare events. A neutrino-driven explosion \citep{Bethe:1990} is the most likely outcome in the case of slowly rotating cores, which are present in the bulk of CCSN progenitors. The emitted GWs could be detected with the advanced ground-based GW detector network, Advanced LIGO (aLIGO)~\citep{TheLIGOScientific:2014jea}, Advanced Virgo (AdV)~\citep{TheVirgo:2014hva} and
KAGRA~\citep{Aso:2013eba}, within $\sim 5$~kpc \citep{Gossan:2016,TargetedSNSearchO12}. Such a galactic event has a rate of about $2-3$ per century \citep{Adams:2013,Rozwadowska:2021}.
For the case of rapidly rotating progenitor cores the result is likely a magneto-rotational explosion, yielding  a more powerful GW signal that could be detected within $50$~kpc and, for some extreme models, up to $5-30$ Mpc \citep{Gossan:2016,TargetedSNSearchO12}. However, only about $1\%$ of the electromagnetically observed events show signatures of fast rotation (broad-lined type Ic SNe \citep{Li:2011b} or events associated with long GRBs \citep{Chapman:2007}) making this possibility a subdominant channel of detection with a galactic event rate of $\sim 10^{-4} \rm{yr}^{-1}$.
%For the results discussed in this work we only consider neutrino-driven CCSN of non-rotating stellar cores.%\tf{This last sentence must be mentioned somewhere (or maybe not) but perhaps here is not the best place.}\mab{MAB:I'm in favor of deleting it as it is rather obvious we are considering neutrino driven CCSN}
Despite the low rates, CCSN are of great scientific interest because they produce complex GW signals which could provide significant clues about the physical processes at work after the gravitational collapse of stellar cores.

@@

All multidimensional numerical simulations show the systematic appearance in time-frequency diagrams (or spectrograms) of a distinct and relatively narrow feature 
%
%{The main feature appearing systematically in the GW spectrum of multidimensional numerical simulations is a strong and relatively narrow \mab{oscillation} 
%
during the post-bounce evolution of the system, with frequency rising 
from about $100$~Hz up to a few kHz (at most) and a typical duration of $0.5-1$~s. This feature has been interpreted as a continuously excited gravity mode (g-mode, see \citep{kokkotas,Friedman:2013} for a definition in this context) of the PNS \citep{murphy:09, mueller:13gw, Cerda:2013, Yakunin:2015, Kuroda:2016, Andresen:2017}. 
In these models the monotonic raise of the frequency of the mode is related to the contraction of the PNS. The {typical} frequencies of {these} modes make them interesting targets for
%promising {source} for 
ground-based GW interferometers.

>>>>

The life of sufficiently massive stars, i.e.~those born with masses between $\sim 8$~M$_\odot$ and $\sim 120$~M$_\odot$, ends with the collapse of {the} iron core under {its} own gravity, leading {to} the formation of a neutron star {(NS)} or a black hole (BH), {and} followed (typically but not necessarily in the BH case) by {a supernova} explosion. Nearby core-collapse supernova (CCSN) explosions are expected to be sources of gravitational waves (GWs) and they could be 
%are one of the main candidates for 
the next great discovery of current ground-based observatories. However, these are relative rare events. A neutrino-driven explosion \citep{Bethe:1990} is the most likely outcome in the case of slowly rotating cores, which are present in the bulk of CCSN progenitors. The emitted GWs could be detected with the advanced ground-based GW detector network, Advanced LIGO (aLIGO)~\citep{TheLIGOScientific:2014jea}, Advanced Virgo (AdV)~\citep{TheVirgo:2014hva} and
KAGRA~\citep{Aso:2013eba}, within $\sim$ \unit[5]{kpc} \citep{Gossan:2016,TargetedSNSearchO12}. Such a galactic event has a rate of about $2-3$ per century \citep{Adams:2013,Rozwadowska:2021}.
For the case of rapidly rotating progenitor cores the result is likely a magneto-rotational explosion, yielding  a more powerful GW signal that could be detected within \unit[50]{kpc} and, for some extreme models, up to \unit[5--30]{Mpc} \citep{Gossan:2016,TargetedSNSearchO12}. However, only about $1\%$ of the electromagnetically observed events show signatures of fast rotation (broad-lined type Ic SNe \citep{Li:2011b} or events associated with long GRBs \citep{Chapman:2007}) making this possibility a subdominant channel of detection with a galactic event rate of $\sim$ \unit[$10^{-4}$]{$\rm {yr}^{-1}$}.
%For the results discussed in this work we only consider neutrino-driven CCSN of non-rotating stellar cores.%\tf{This last sentence must be mentioned somewhere (or maybe not) but perhaps here is not the best place.}\mab{MAB:I'm in favor of deleting it as it is rather obvious we are considering neutrino driven CCSN}
Despite the low rates, CCSN are of great scientific interest because they produce complex GW signals which could provide significant clues about the physical processes at work after the gravitational collapse of stellar cores.

@@

All multidimensional numerical simulations show the systematic appearance in time-frequency diagrams (or spectrograms) of a distinct and relatively narrow feature 
%
%{The main feature appearing systematically in the GW spectrum of multidimensional numerical simulations is a strong and relatively narrow \mab{oscillation} 
%
during the post-bounce evolution of the system, with frequency rising 
from about \unit[100]{Hz} up to a few kHz (at most) and a typical duration of \unit[0.5 -- 1]{s}. This feature has been interpreted as a continuously excited gravity mode (g-mode, see \citep{kokkotas,Friedman:2013} for a definition in this context) of the PNS \citep{murphy:09, mueller:13gw, Cerda:2013, Yakunin:2015, Kuroda:2016, Andresen:2017}. 
In these models the monotonic raise of the frequency of the mode is related to the contraction of the PNS. The {typical} frequencies of {these} modes make them interesting targets for
%promising {source} for 
ground-based GW interferometers.

### 2020-11-10 methods.tex
Pablo s comemnts answer
prep version to be circulated in LVK

\begin{figure}[t]
 \centering
 \includegraphics[width=0.45\textwidth]{plots/model}
 \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ from our 25 1D simulations of the model set. The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering the 18 simulation data points. \tf{18 or 25?} \tf{We could remove the legend as only one code is used.} } 
 \label{fig:LMVAR}
\end{figure}

We next outline our strategy for estimating the time evolution of $r(t)$
%ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ (in units of solar mass and km) 
from the observation of the $\mbox{}^2g_2$ oscillation mode in the GW detector data.
%An integral part of this strategy is the universal relations that relate the
%characteristic frequency of the PNS oscillation $f$, $g$ and $p$ modes with the mass
%and the radius of the PNS, the shock radius and the total mass inside the shock as
%demonstrated in \cite{Torres:2019b}.
To build the model of the ratio $r$ as a function of the frequency $f$ we use the 
1D simulations of the {\it model set}. Figure~\ref{fig:LMVAR}
shows the data for the $25$ numerical simulations. 
%As identified by \cite{Torres:2019b}, the only systematic deviation from a single universal relation is the numerical code used in the simulations. 
%To avoid any systematic effect, we only use the $18$ simulations performed with the {\sc AENUS-ALCAR}
%code, which is the same code that was used in our test set.
%\mab{We defer a discussion of the possible consequences of this choice in Section \ref{sec:conclusion}.}
%The consequences of this choice are discussed in the conclusions.
Using these data we parametrize the discretized ratio $r_i$ with a cubic polynomial
regression with heteroscedastic errors
\begin{equation}
\label{eq:model1}
r_i=\beta_1 f_i + \beta_2 f_i^2 +\beta_3 f_i^3 + \epsilon_i\,,
\end{equation}
where $\epsilon_i$ are assumed to be independent zero-mean Gaussian errors with
variances $\sigma_i^2$ that increase with frequency $f_i$. The model for frequency-dependent
variances is
\begin{equation}
\log \sigma_i=\alpha_0+ \alpha_1 f_i + \alpha_2 f_i^2 + \delta_i\,,
\end{equation}
with independent and identically zero-mean Gaussian errors $\delta_i$. The R-package \texttt{lmvar}
\cite{lmvar:2019} that implements a maximum likelihood approach was used to fit the model.

@@

{We use this model to infer the properties of the simulations in the 
{\it test set} discussed in Section \ref{sec:simulations}.}
To describe the method we focus on the GW signal
of model {\texttt s20S}, originally
sampled at 16384 Hz but downsampled at 4096 Hz.
A spectrogram of this signal is shown in Figure \ref{fig:spectrogram} based on
autoregressive estimates~\cite{BrockwellPeterJ1991Tsta} of the local spectra for successive time intervals of 
length 200 with a 90\% overlap.
The dominant emission mode corresponds to the PNS oscillation $\mbox{}^2 g_2$-mode. We have
developed a time-frequency method to track the ridge $m(t) $ in the spectrogram,
taking into account that it is monotonically increasing with time. 
This is a property of the $\mbox{}^2 g_2$-mode whose frequency   
increases as the PNS becomes more massive and compact.
Starting from either the left- or right-most column of the time-frequency matrix
we identify and trace the sequence of amplitude peaks within a certain frequency
band given the monotonicity constraint. Specific details about the reconstruction of the $\mbox{}^2 g_2$-mode ridge 
are provided in Appendix \ref{app:gmode}.

@@

We collect the instantaneous frequency $f(t_i)$ corresponding to the ridge $m(t_i)$ for
the midpoint $t_i$ of each local time interval of the spectrogram and interpolate $f(t)$
for values in between $t_i$. We then use our model given by Eq.~\eqref{eq:model1} to obtain
estimates of the time evolution of the ratio together with 95\% confidence intervals.
An example is given in Figure \ref{fig:ratio} where the red triangles are the point estimates and
the grey bands represent 95\% confidence bands. The size of the red triangles is proportional to the magnitude of the $\mbox{}^2 g_2$-mode estimates.  Note that as the frequency of the $\mbox{}^2 g_2$-mode becomes higher our estimates show more uncertainty (bigger intervals) because our model allows for heterogeneous variance. Ratio values
computed using the mass and radius values obtained from the simulation code (i.e.~the true values)
are shown in black. In this example of a noise-free GW signal the coverage of our 95\% confidence band is 100\%
of the true values. In the next section we investigate the performance of the reconstruction of $r(t)$ when the GW
signal is embedded in noise. We also note that despite we have only explicitely shown results for the GW signal of the {\texttt s20S} model the same conclusions hold for any of the other waveforms of our test set.

\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth]{plots/ratio}
 \caption{Comparison of the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$ estimated from the $\mbox{}^2 g_2$-mode of the {\texttt s20S} signal (shown by open triangles and by the 95\% confidence belt in grey) against the value derived from the PNS mass and radius given by the simulation code (shown by filled black circles).  The ratios are represented proportionally to the magnitude of the $\mbox{}^2 g_2$-mode estimates.}
 \label{fig:ratio}
\end{figure}

>>>>

\begin{figure}[t]
 \centering
 \includegraphics[width=0.45\textwidth]{plots/model}
 \caption{Ratio $M_{\rm PNS}/R_{\rm PNS}^2$ from our 18 1D simulations of the model set. The solid line is the maximum likelihood estimate of heteroscedastic cubic model with 95\% confidence bands (dashed lines) considering the 18 simulation data points.} 
 \label{fig:LMVAR}
\end{figure}

We next outline our strategy for estimating the time evolution of $r(t)$
%ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ (in units of solar mass and km) 
from the observation of the $\mbox{}^2g_2$ oscillation mode in the GW detector data.
%An integral part of this strategy is the universal relations that relate the
%characteristic frequency of the PNS oscillation $f$, $g$ and $p$ modes with the mass
%and the radius of the PNS, the shock radius and the total mass inside the shock as
%demonstrated in \cite{Torres:2019b}.
To build the model of the ratio $r$ as a function of the frequency $f$ we use the 
1D simulations of the {\it model set}. Figure~\ref{fig:LMVAR}
shows the data for the $18$ numerical simulations. 
%As identified by \cite{Torres:2019b}, the only systematic deviation from a single universal relation is the numerical code used in the simulations. 
%To avoid any systematic effect, we only use the $18$ simulations performed with the {\sc AENUS-ALCAR}
%code, which is the same code that was used in our test set.
%\mab{We defer a discussion of the possible consequences of this choice in Section \ref{sec:conclusion}.}
%The consequences of this choice are discussed in the conclusions.
Using these data we parametrize the discretized ratio $r_i$ with a cubic polynomial
regression with heteroscedastic errors
\begin{equation}
\label{eq:model1}
r_i=\beta_1 f_i + \beta_2 f_i^2 +\beta_3 f_i^3 + \epsilon_i\,,
\end{equation}
where $\epsilon_i$ are assumed to be independent zero-mean Gaussian errors with
variances $\sigma_i^2$ that increase with frequency $f_i$. The model for frequency-dependent
variances is
\begin{equation}
\log \sigma_i=\alpha_0+ \alpha_1 f_i + \alpha_2 f_i^2 + \delta_i\,,
\end{equation}
with independent and identically zero-mean Gaussian errors $\delta_i$. The R-package \texttt{lmvar}
\cite{lmvar:2019} that implements a maximum likelihood approach was used to fit the model.

@@

{We use this model to infer the properties of the simulations in the 
{\it test set} discussed in Section \ref{sec:simulations}.}
To describe the method we focus on the GW signal
of model {\texttt s20S}, originally
sampled at \unit[16384]{Hz} but downsampled at \unit[4096]{Hz}.
A spectrogram of this signal is shown in Figure \ref{fig:spectrogram} based on
autoregressive estimates~\cite{BrockwellPeterJ1991Tsta} of the local spectra for successive time intervals of 
length 200 with a 90\% overlap.
The dominant emission mode corresponds to the PNS oscillation $\mbox{}^2 g_2$-mode. We have
developed a time-frequency method to track the ridge $m(t) $ in the spectrogram,
taking into account that it is monotonically increasing with time. 
This is a property of the $\mbox{}^2 g_2$-mode whose frequency   
increases as the PNS becomes more massive and compact.
Starting from either the left- or right-most column of the time-frequency matrix
we identify and trace the sequence of amplitude peaks within a certain frequency
band given the monotonicity constraint. Specific details about the reconstruction of the $\mbox{}^2 g_2$-mode ridge 
are provided in Appendix \ref{app:gmode}.

@@

We collect the instantaneous frequency $f(t_i)$ corresponding to the ridge $m(t_i)$ for
the midpoint $t_i$ of each local time interval of the spectrogram and interpolate $f(t)$
for values in between $t_i$. We then use our model given by Eq.~\eqref{eq:model1} to obtain
estimates of the time evolution of the ratio together with 95\% confidence intervals.
An example is given in Figure \ref{fig:ratio} where the red triangles are the point estimates and
the grey bands represent 95\% confidence bands. The size of the red triangles is proportional to the magnitude of the $\mbox{}^2 g_2$-mode frequency estimates.
Note that as the frequency of the $\mbox{}^2 g_2$-mode becomes higher our estimates show more uncertainty (bigger intervals) because our model allows for heterogeneous variance. Ratio values
computed using the mass and radius values obtained from the simulation code (i.e.~the true values)
are shown in black. In this example of a noise-free GW signal the coverage of our 95\% confidence band is 100\%
of the true values. In the next section we investigate the performance of the reconstruction of $r(t)$ when the GW
signal is embedded in noise. We also note that despite we have only explicitely shown results for the GW signal of the {\texttt s20S} model the same conclusions hold for any of the other waveforms of our test set.

\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth]{plots/ratio}
 \caption{Comparison of the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$ estimated from the $\mbox{}^2 g_2$-mode of the {\texttt s20S} signal (shown by open triangles and by the 95\% confidence belt in grey) against the value derived from the PNS mass and radius given by the simulation code (shown by filled black circles). The size of the triangles are represented proportionally to the magnitude of the $\mbox{}^2 g_2$-mode frequency estimates.}
 \label{fig:ratio}
\end{figure}

### 2020-11-10 results.tex
prep version to be circulated in LVK

For each of the noise realizations we reconstruct the ratio time series {$r_i$} of length $N$ starting from the left side of the spectrogram and constraining the beginning of the track to be smaller than \unit[200]{Hz}. \mab{This value is chosen using the information on the initial mode frequency from the simulations.}
The reconstructed ratio is then compared to the ``true'' ratio {$r_i^0$} derived from the PNS mass and radius computed from the {\texttt s20S} simulation. The top panel of Fig.~\ref{fig:s20results} shows the distribution of the fraction of the ratio values {$r_i^0$} that fall within the 95\% confidence interval of {$r_i$} as a function of the distance to the source. This quantity, hence, gives information about the {\it coverage} of the reconstructed ratio. The coverage takes maximum values when the source is located within a few kpc and then decreases with the distance.

To better quantify how well we reconstruct the ratio we also consider  the mean of the relative error of $r_i$ along 
the track of the spectrogram, $\Delta$,  
\begin{equation}
\Delta=\frac{1}{N}\sum_1^N\frac{|r_i-r_i^0|}{r_i^0}\,.
\end{equation}
The values of $\Delta$ for each of the 100 noise realizations are shown as a function of the distance
in the bottom panel of Fig.~\ref{fig:s20results}. For a source located up to $\sim$\unit[9]{kpc} the relative error
remains smaller than 20\%. At closer distances $\Delta$ is smaller but it does not vanish. This reflects the approximate nature of the model used for $r$. It is nevertheless remarkable that, on average, one can reconstruct the ratio time series with a good
precision up to distances of $\sim$ 9 kpc (for this particular waveform) with a coverage value
larger than 80\%. We note that there are a few noise realizations for distances below 9 kpc for which
$\Delta$ takes large values, indicating that the method fails to accurately reconstruct the ratio in those cases. The main cause of failure is the split of the ridge in few blobs when the signal becomes weaker and weaker.
%\tf{Do we know why that is the case? What is intrinsically different from one noise realisation to another that makes the method more prone to failure? I think we should provide some explanation here, if we have it.}

@@

We have tested that the method does not depend on the specific features of the waveform of model {\texttt s20S} by repeating the procedure for the remaining seven waveforms of the {\it test set} described in Section \ref{sec:simulations} covering
a large range of progenitor masses. Figure \ref{fig:aLIGO_cov_allwvf} shows that apart from model {\texttt s11} and to a lesser extent model {\texttt s20S}, the ratio is well reconstructed for all waveforms up to a distance of $\sim$ 15kpc. In an
effort to better determine the maximal distance of the source at which we can reconstruct the ratio we have run 100 simulations without injecting a signal and have measured the corresponding coverage for the reconstructed ratios.
The median of the coverage as well as the band defined by the $\mathrm{5^{th}}$ and $\mathrm{95^{th}}$ percentiles are shown in Figure \ref{fig:aLIGO_cov_allwvf}.
The noise only median value is identically zero in this case. However, note that it could be different
from zero because
the g-mode reconstruction algorithm is looking for a continuously  increasing frequency track
in the spectrogram, starting between 0 and 200 Hz, where we expect the GW signal to be.
This is enhancing the probability of overlap. This effect explains why certain values of overlap can reach
values as high as 80\% even when no signal is added to the noise. \tf{I don't understand this last comment.} \mab{Toni: is that clearer?}

@@

We perform the same analysis using the design sensitivity curve of AdV and expected sensitivity curves for third-generation 
GW detectors. The results for the former are reported in Table \ref{tab:results} and are discussed below. We focus now on third-generation detectors, presenting our findings in Table \ref{tab:results} and in Fig.~\ref{fig:s20--SFHo_all3G}. In Europe the Einstein Telescope project proposes to host in a 10-km equilateral triangle configuration three low-power, low-frequency, cryogenic interferometers as well as three high-power, high-frequency interferometers. Three sensitivity curves, ET-B, ET-C and ET-D corresponding to different options and stages of the project~\cite{Hild_2011} are considered in our study. %(cf.~Fig.~\ref{fig:spectrum}).
The US based project Cosmic Explorer~\cite{reitze2019cosmic} is expecting to reach its design
sensitivity circa 2040 through two phases labeled CE1 and CE2. %Their corresponding sensitivity curves are also shown in Fig.~\ref{fig:spectrum}.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\textwidth]{plots/s20--SFHo_all3G}
  \caption{Median of the relative error $\Delta$ as a function of the  distance to the source for the CCSN waveform model {\texttt s20S} embedded in third-generation detector noise. \pcd{[PCD: Should we show the corresponding blue ranges for the pure noise case?]}\mab{MAB: We would need to show 5 noise ranges. I haven't tried what it looks like. Could give a try}}
  \label{fig:s20--SFHo_all3G}
\end{figure}

Figure \ref{fig:s20--SFHo_all3G} displays $\Delta$ as a function of the source distance for the five third-generation detector configurations we analyze. As before, we use the {\texttt s20S} waveform model as a reference case.  This figure shows that,  overall, the ratio is well reconstructed up to distances in the range \unit[100--200]{kpc} which represents an order of magnitude improvement with respect to aLIGO and AdV. We also note that the results for the various Einstein Telescope configurations lay in between those of the two Cosmic Explorer designs. Moreover,  the detectability prospects for the former depend weakly on the detector configuration while the arrangement of CE2 yields better results than CE1. These results are confirmed for all other waveforms of our sample except for model {\texttt s25} for which the maximal distance reach in CE2 is significantly lower than CE1, as shown in Fig.~\ref{fig:distances}. This is partly due to the small variation of the reconstruction quality to the distance of the source making the estimation of $d_r$ rather uncertain for this particular waveform. \tf{Is this really understood? What's particular in this waveform?}

@@

All results for both second-generation and third-generation detectors are summarized in Table \ref{tab:results} and Figure
\ref{fig:distances}. Table \ref{tab:results} reports the source distances $d_r$ in kpc at which the median of the coverage is lower than 95\% of the noise only values for aLIGO, AdV, and different configurations of third-generation detectors. This same information is displayed in Fig.~\ref{fig:distances}. We have checked that using either the median of the coverage or the median of $\Delta$ yields similar results for the distance. On the other hand, the quality of the ratio reconstruction and, thus, of the distance range, depends on the signal-to-noise ratio, expressed in Table \ref{tab:results} by $d_{\rm det}$. The numbers reported on the table for $d_r$ are an estimate of the order of magnitude of the maximal distance of the source  at which a
reconstruction of the ratio could be possible with current and planned GW detectors. We also provide upper limits for $d_{\rm det}$ by taking into account the detector antenna response in our simulations and assuming that the source is optimally oriented with a matched filter signal-to-noise ratio of 13. Table \ref{tab:results} shows that the results for the Advanced Virgo detector at design sensitivity are very similar to those of aLIGO, despite the differences in detector sensitivity. It is remarkable that for third-generation detectors the PNS surface gravity  could be reconstructed for sources located up to a few hundreds of kpc. It is nevertheless important to note the rather wide range in distances we obtain for the different waveforms of our {\it test set} that probe a large range of progenitor masses. We do not find any correlation between either the mass of the progenitor nor the EOS with $d_r$.

>>>>

For each of the noise realizations we reconstruct the track ridge $m(t_i)$ starting from the left side of the spectrogram and constraining the beginning of the track to be smaller than \unit[200]{Hz}. This value is chosen using the information on the initial mode frequency from the simulations. We derive the ratio time series $r_i$ of length $N$ which is then compared to the ``true'' ratio {$r_i^0$} derived from the PNS mass and radius computed from the {\texttt s20S} simulation. The top panel of Fig.~\ref{fig:s20results} shows the distribution of the fraction of the ratio values {$r_i^0$} that fall within the 95\% confidence interval of {$r_i$} as a function of the distance to the source. This quantity, hence, gives information about the {\it coverage} of the reconstructed ratio. The coverage takes maximum values when the source is located within a few kpc and then decreases with the distance.

To better quantify how well we reconstruct the ratio we also consider  the mean of the relative error of $r_i$ along 
the track of the spectrogram, $\Delta$,  
\begin{equation}
\Delta=\frac{1}{N}\sum_1^N\frac{|r_i-r_i^0|}{r_i^0}\,.
\end{equation}
The values of $\Delta$ for each of the 100 noise realizations are shown as a function of the distance
in the bottom panel of Fig.~\ref{fig:s20results}. For a source located up to $\sim$\unit[9]{kpc} the relative error
remains smaller than 20\%. At closer distances $\Delta$ is smaller but it does not vanish. This reflects the approximate nature of the model used for $r$. It is nevertheless remarkable that, on average, one can reconstruct the ratio time series with a good
precision up to distances of $\sim$ \unit[9]{kpc} (for this particular waveform) with a coverage value
larger than 80\%. We note that there are a few noise realizations for distances below \unit[9]{kpc} for which
$\Delta$ takes large values, indicating that the method fails to accurately reconstruct the ratio in those cases. The main cause of failure is the split of the ridge in few blobs when the signal becomes weaker and weaker.
%\tf{Do we know why that is the case? What is intrinsically different from one noise realisation to another that makes the method more prone to failure? I think we should provide some explanation here, if we have it.}

@@

We have tested that the method does not depend on the specific features of the waveform of model {\texttt s20S} by repeating the procedure for the remaining seven waveforms of the {\it test set} described in Section \ref{sec:simulations} covering
a large range of progenitor masses. Figure \ref{fig:aLIGO_cov_allwvf} shows that apart from model {\texttt s11} and to a lesser extent model {\texttt s20S}, the ratio is well reconstructed for all waveforms up to a distance of $\sim$ 15kpc. In an
effort to better determine the maximal distance of the source at which we can reconstruct the ratio we have run 100 simulations without injecting a signal and have measured the corresponding coverage for the reconstructed ratios.
The median of the coverage as well as the band defined by the $\mathrm{5^{th}}$ and $\mathrm{95^{th}}$ percentiles are shown in Figure \ref{fig:aLIGO_cov_allwvf}.
The noise only median value is identically zero in this case. However, note that it could be different
from zero because
the g-mode reconstruction algorithm is looking for a continuously  increasing frequency track
in the spectrogram, starting between 0 and 200 Hz, where we expect the GW signal to be.
This is enhancing the probability of overlap. This effect explains why certain values of overlap can reach
values as high as 80\% even when no signal is added to the noise.
%\tf{I don't understand this last comment.} \mab{Toni: is that clearer?}

@@

We perform the same analysis using the design sensitivity curve of AdV and expected sensitivity curves for third-generation 
GW detectors. The results for the former are reported in Table \ref{tab:results} and are discussed below. We focus now on third-generation detectors, presenting our findings in Table \ref{tab:results} and in Fig.~\ref{fig:s20--SFHo_all3G}. In Europe the Einstein Telescope project proposes to host in a 10-km equilateral triangle configuration three low-power, low-frequency, cryogenic interferometers as well as three high-power, high-frequency interferometers. Three sensitivity curves, ET\_B, ET\_C and ET\_D corresponding to different options and stages of the project~\cite{Hild_2011} are considered in our study. %(cf.~Fig.~\ref{fig:spectrum}).
The US based project Cosmic Explorer~\cite{reitze2019cosmic} is expecting to reach its design
sensitivity circa 2040 through two phases labeled CE1 and CE2. %Their corresponding sensitivity curves are also shown in Fig.~\ref{fig:spectrum}.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\textwidth]{plots/s20--SFHo_all3G}
  \caption{Data points (``+'') show the median of the relative error $\Delta$ as a function of the  distance to the source for the CCSN waveform model {\texttt s20S} embedded in third-generation detector noise. The solid lines are smoothing splines. The blue band boundaries are given by the $\mathrm{5^{th}}$ and $\mathrm{95^{th}}$ percentiles of coverage in absence of any signal for ET\_D noise realisations.}
  \label{fig:s20--SFHo_all3G}
\end{figure}

Figure \ref{fig:s20--SFHo_all3G} displays $\Delta$ as a function of the source distance for the five third-generation detector configurations we analyze. As before, we use the {\texttt s20S} waveform model as a reference case.  This figure shows that,  overall, the ratio is well reconstructed up to distances in the range \unit[100--200]{kpc} which represents an order of magnitude improvement with respect to aLIGO and AdV. We also note that the results for the various Einstein Telescope configurations lay in between those of the two Cosmic Explorer designs. Moreover,  the detectability prospects for the former depend weakly on the detector configuration while the arrangement of CE2 yields better results than CE1. These results are confirmed for all other waveforms of our sample except for model {\texttt s25} for which the maximal distance reach in CE2 is significantly lower than CE1, as shown in Fig.~\ref{fig:distances}. This is partly due to the small variation of the reconstruction quality to the distance of the source making the estimation of $d_r$ rather uncertain for this particular waveform. %\tf{Is this really understood? What's particular in this waveform?}

@@

All results for both second-generation and third-generation detectors are summarized in Table \ref{tab:results} and Figure
\ref{fig:distances}. Table \ref{tab:results} reports the source distances $d_r$ in kpc at which the median of the coverage is lower than 95\% of the noise only values for aLIGO, AdV, and different configurations of third-generation detectors. This same information is displayed in Fig.~\ref{fig:distances}. We have checked that using either the median of the coverage or the median of $\Delta$ yields similar results for the distance. On the other hand, the quality of the ratio reconstruction and, thus, of the distance range, depends on the signal-to-noise ratio, expressed in Table \ref{tab:results} by $d_{\rm det}$. The numbers reported on the table for $d_r$ are an estimate of the order of magnitude of the maximal distance of the source  at which a
reconstruction of the ratio could be possible with current and planned GW detectors. We also provide upper limits for $d_{\rm det}$ by taking into account the detector antenna response in our simulations and assuming that the source is optimally oriented with a matched filter signal-to-noise ratio of 13. Table \ref{tab:results} shows that the results for the AdV detector at design sensitivity are very similar to those of aLIGO, despite the differences in detector sensitivity. It is remarkable that for third-generation detectors the PNS surface gravity  could be reconstructed for sources located up to a few hundreds of kpc. It is nevertheless important to note the rather wide range in distances we obtain for the different waveforms of our {\it test set} that probe a large range of progenitor masses. We do not find any correlation between either the mass of the progenitor nor the EOS with $d_r$.
